---
title: "Practical Machine Learning Project"
output: github_document
---

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. The goal of your project is to predict the manner in which they did the exercise. This is the "classe" variable in the training set. You may use any of the other variables to predict with.

```{r}
library(caret)
library(randomForest)
library(dplyr)
```

```{r}
setwd("~/Documents/Data Science Specialization/Practical Machine Learning")
train <- read.csv("pml-training.csv")
test <- read.csv("pml-testing.csv")
```

First we must partition the training dataset to test our model. I've changed the classe variable from "character" to "factor" as it will be important once we build our model.
```{r}
set.seed(1223)
trainingdata <- createDataPartition(train$classe, p = .8, list = FALSE)
Training <- train[trainingdata, ]
Validation <- train[-trainingdata, ]
Training$classe <- as.factor(Training$classe)
Validation$classe <- as.factor(Validation$classe)
```

Now we clean the data to get rid of unhelpful columns (NA columns, columns with mostly spaces etc.) and descriptive columns.
```{r}
Training <- Training[, colSums(is.na(Training)) == 0]
Training <- Training[, colSums(Training == "") == 0]
Training <- Training[, -c(1:7)]
```

Next, we build the model using a random forest algorithm.
```{r}
model <- randomForest(classe ~ ., data = Training, importance = TRUE, ntrees = 15)
```

Then we test it against the training set and use the validation partition to cross validate.
```{r}
self_train <- predict(model, Training)
confusionMatrix(self_train, Training$classe)

validation <- predict(model, Validation)
confusionMatrix(validation, Validation$classe)
```
The accuracy is at 0.995 so our out of sample error should be at 0.005. Since that error rate is very low, we can say the model performs considerably well.

Finally, we use our model on the test data set.
```{r}
test_final <- predict(model, test)
test_final
```
